---
title: "STAT 466 HW 11"
author: "Brody Anderson"
date: "2024-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R2jags)
library(survival)
set.seed(1234)
```


## Question 1

Similar to the experiment in Example 4.8, Ku et al. (1972) reports on an experiment with a different lubricant called O-67-22. Table 7.31 presents the resulting lifetimes. Is σ2 for O-67-22 similar to that found in Exam- ple 4.8? Is μ for this lubricant different from that for the other lubricant? Also fit the data from both experiments in one model using an appropriate covariate and assuming a common σ2 for the two lubricants. Based on the combined analysis, is one lubricant better than the other?


```{r Q1}
# Table 4.5 for lubricant O-64-2
# Table 7.31 for lubricant O-67-22


```



## Question 2

```{r Q2}

```



## Question 3

7.22 The inverse power or simply power relationship is another model relating lifetime to an accelerating stress, such as elevated voltage. For a positive stress variable V , the nominal lifetime takes the form A/V γ1 . For Weibull lifetimes, we can express the characteristic life ψ (third parameterization of the Weibull distribution in Appendix B) as log(ψ) = γ0 +γ1[− log(V )].

The data displayed in Table 7.42 from Nelson (1990) are breakdown times of an insulating fluid at seven high voltages.

(a) Analyze the lifetime data using the Weibull power relationship model.
(b) Predict the mean lifetime and lifetime distribution at 20 kV.

```{r Q3}

```


